Adolescent well-being in the digital age

### 2021-12-17 index.qmd
more edits

---
title: "Adolescent well-being in the digital age"
author: "Matti Vuorre"
date: "`r Sys.time()`"
execute:
  eval: true
  error: true
  cache: true
  warning: false
  message: false
format: 
  html:
    monofont: Fira Code
    font-size: 1.0em
    linestretch: 1.4
    code-fold: true
    code-summary: "Show code"
    code-tools: true
    code-link: true
toc: true
toc-depth: 4
toc-title: Contents
number-sections: false
crossref:
  fig-prefix: Figure
  tbl-prefix: Table
---

@@

# Format tables
kable_custom <- function(x) {
  kable_classic(
    kable_input = x,
    html_font = Font,
    font_size = 13,
    lightable_options = "striped",
    position = "center",
    full_width = FALSE
  )
}

@@

```{r fit-model-1}

fits <- dat %>%
  # Could nest but need this to include group key in nested tibble
  group_by(outcome) %>%
  summarise(data = list(cur_data_all()))

@@

```{r model-1-data-summary}
#| column: body-outset
#| tbl-cap: "Summaries of data included in Model 1"
#| label: tbl-m1-data
msum <- function(x) {
  tibble(
    regions = ngrps(x)$region,
    countries = ngrps(x)$country,
    `year/country` = x$data %>%
      distinct(country, year) %>%
      count(country) %>%
      pull(n) %>%
      median(),
    nobs_internet = standata(x)$N_internet,
    nobs_outcome = standata(x)$N_val,
    outcome_range = str_glue(
      "{number(min(x$data$val), .01)} - {number(max(x$data$val), .01)}"
    )
  )
}
fits %>%
  mutate(summary = map(fit1, msum)) %>%
  select(outcome, summary) %>%
  unnest(summary) %>%
  kbl(caption = "Model 1 data summary") %>%
  kable_custom() %>% 
  row_spec(0, bold = TRUE)
```

@@

```{r ppcheck-density}

#| column: page
#| fig-cap: "Graphical posterior predictive check by region (Model 1). Thin red lines indicate model replicates, whereas the blue lines indicate data; the model is mostly in smooth agreement with data, but has greater discrepancies in some regions (e.g. Anglo Depression)."
#| label: fig-ppcheck-1
ppcheck_1 <- fits %>%
  group_by(outcome) %>%
  mutate(
    out = map(
      fit1,
      ~ pp_check(
        .x,
        ndraws = 10,
        type = "dens_overlay_grouped",
        group = "region",
        resp = "val",
        newdata = .x$data
      )[[1]]
    )
  ) %>%
  ungroup() %>% 
  select(outcome, out) %>%
  unnest(out)

ppcheck_1 %>%
  ggplot(aes(value, col = is_y, alpha = is_y, group = rep_id)) +
  scale_color_brewer(palette = "Set1") +
  scale_alpha_manual(values = c(.2, 1)) +
  stat_density(
    geom = "line",
    size = .75,
    position = position_identity()
  ) +
  labs(x = "Value", y = "Density") +
  facet_grid2(
    outcome ~ group,
    scales = "free", independent = "x",
    labeller = as_labeller(function(x) str_replace(x, "_", "\n"))
  ) +
  theme(
    legend.position = "none",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
```

@@

```{r calculate-simple-effects}

#' Posterior draws at average and region levels
#'
#' @param x brmsfit (model 1/2)
#' @param p name of parameter
#'
#' @return a long tibble posterior draws of p for every region (including the average, so useful for e.g. forest plots)
pdraws <- function(x, p = "val_year") {
  bind_cols(
    tibble("Average" = fixef(x, summary = FALSE)[, p]),
    as_tibble(coef(x, summary = FALSE)$region[, , p])
  ) %>%
    pivot_longer(everything(), names_to = "region")
}

@@

```{r shrinkage}

#| column: screen-inset
#| fig-width: 14
#| fig-height: 9
#| fig-cap: "Scatterplots of multilevel and independent model coefficients"
#| label: fig-trend-shrinkage

@@

```{r model21-compare}

#| results: hold
#| tbl-cap: "Model 1 and 2 proportions variance explained by outcome"
#| label: tbl-m12-r2
fits %>%
  select(outcome, fit1, fit2) %>% 
  pivot_longer(c(fit1, fit2)) %>% 
  mutate(
    r2 = map(
      value,
      ~ bayes_R2(., resp = "val") %>%
        as.data.frame() %>%
        mutate(across(everything(), ~ percent(., .1))) %>%
        transmute(result = str_glue("{Estimate} [{Q2.5}, {Q97.5}]"))
    )
  ) %>%
  select(-value) %>%
  unnest(r2) %>% 
  pivot_wider(
    names_from = name,
    values_from = result,
    names_prefix = "Model "
  ) %>%
  mutate(outcome = str_replace(outcome, "_", " ")) %>%
  kbl(caption = "Model 1 and 2 R-squared values") %>%
  kable_custom()
```

@@

```{r model2-calculate-coefficients}

p_draws <- fits %>%
  mutate(pdraws = map(fit2, ~pdraws(.x, "i1_cw"))) %>%
  select(outcome, pdraws) %>%
  unnest(pdraws)

@@

```{r model2-fitted-values}

#| column: screen-inset
#| fig-width: 14
#| fig-height: 9
#| fig-cap: "Counterfactual plot of internet associations"
#| label: fig-fitted-internet

@@

```{r}
#| fig.width: 6
#| fig.height: 4
#| fig-cap: "Between-country associations"
#| label: fig-cb1

@@

  newy <- bind_cols(
    newx,
    as.data.frame(
      fitted(
        x, 
        newdata = newx, 
        re_formula = NULL
        )
    )
  ) %>%
    select(-outcome)
  newy
}
tmp <- fits %>% 
  mutate(out = map(fit2, fitted_cb1)) %>% 
  select(outcome, out) %>%
  unnest(out)

@@

tmp %>% 
  ggplot(aes(i1_cb, val)) +
  scale_x_continuous(
    "Between-country centered internet adoption",
    labels = ~percent(.)
  ) +
  scale_y_continuous(
    "Value",
    labels = ~percent(./100)
  ) +
  geom_point(
    shape = 1, size = 1, alpha = .75,
    aes(col = region)
  ) +
  geom_ribbon(
    aes(ymin = Q2.5, ymax = Q97.5),
    fill = "dodgerblue2",
    alpha = .2
    ) +
  geom_line(aes(y = Estimate), col = "dodgerblue4") +
  geom_text(
    data = cb1_pars,
    aes(x = -Inf, y = Inf, label = str_glue("{mean} [{q2.5}, {q97.5}]")),
    vjust = 1.3, hjust = -0.025, family = Font
  ) +
  facet_wrap(
    "outcome",
    scales = "free_y", nrow = 2,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(
    aspect.ratio = 1,
    axis.title.y = element_blank()
  )
```

@@

```{r model1-demographics}

h_gwp <- c(
  `15 to 19` =
    "val_year + val_year:age1 * 1 = 0",
  `20 to 24` =
    "val_year + val_year:age1 * -1 = 0",
  `Female` =
    "val_year + val_year:sex1 * 1 = 0",
  `Male` =
    "val_year + val_year:sex1 * -1 = 0"
)
h_gbd <- c(
  `10 to 14` =
    "val_year + val_year:age1 * 1 + val_year:age2 * 0 = 0",
  `15 to 19` =
    "val_year + val_year:age1 * 0 + val_year:age2 * 1 = 0",
  `20 to 24` =
    "val_year + val_year:age1 * -1 + val_year:age2 * -1 = 0",
  `Female` =
    "val_year + val_year:sex1 * 1 = 0",
  `Male` =
    "val_year + val_year:sex1 * -1 = 0"
)
#' Calculate posterior draws for simple effects (by sex and age) for regions and average
#'
#' @param x brmsfit model 2
#' @param h a brmshypothesis
#'
#' @return a long tibble of draws of posterior answers to hypothesis
simple_effects_samples <- function(x, h) {
  # Calculate hypotheses at average level
  average_h <- hypothesis(
    x, h
  )
  # Give hypotheses appropriate names
  average <- average_h$samples
  names(average) <- interaction(
    "Average", average_h$hypothesis$Hypothesis,
    sep = "+"
  )
  # Pivot to long format, note use of dtplyr
  average <- average %>%
    lazy_dt() %>%
    pivot_longer(everything()) %>%
    separate(name, into = c("region", "Hypothesis"), sep = "\\+") %>%
    as_tibble()

@@

```{r model2-demographics}

h_gwp <- c(
  `15 to 19` =
    "i1_cw + age1:i1_cw * 1 = 0",
  `20 to 24` =
    "i1_cw + age1:i1_cw * -1 = 0",
  `Female` =
    "i1_cw + sex1:i1_cw * 1 = 0",
  `Male` =
    "i1_cw + sex1:i1_cw * -1 = 0"
)
h_gbd <- c(
  `10 to 14` =
    "i1_cw + age1:i1_cw * 1 + age2:i1_cw * 0 = 0",
  `15 to 19` =
    "i1_cw + age1:i1_cw * 0 + age2:i1_cw * 1 = 0",
  `20 to 24` =
    "i1_cw + age1:i1_cw * -1 + age2:i1_cw * -1 = 0",
  `Female` =
    "i1_cw + sex1:i1_cw * 1 = 0",
  `Male` =
    "i1_cw + sex1:i1_cw * -1 = 0"
)

>>>>

---
title: "Adolescent well-being in the digital age"
author: "Matti Vuorre"
execute:
  enabled: true
  cache: true
  warning: false
  message: false
format: 
  html:
    monofont: Fira Code
    font-size: 1.0em
    code-fold: true
toc: true
toc-depth: 4
crossref:
  fig-prefix: Figure
  tbl-prefix: Table
---

@@

#' Custom kableExtra formatting
#'
#' @param a table from kbl()
#'
#' @return a table
kable_custom <- function(x) {
  kable_classic(
    kable_input = x,
    html_font = Font,
    font_size = 13,
    lightable_options = "striped",
    position = "center",
    full_width = FALSE
  )
}

@@

```{r fit-model-1}
fits <- dat %>%
  # Could nest but need this to include group key in nested tibble
  group_by(outcome) %>%
  summarise(data = list(cur_data_all()))

@@

```{r model-1-data-summary}
#| column: body-outset
#| tbl-cap: "Summaries of data included in Model 1"
#| label: tbl-m1-data

#' Model summary
#'
#' @param x a brmsfit
#'
#' @return A tibble with information on data in the model
msum <- function(x) {
  tibble(
    regions = ngrps(x)$region,
    countries = ngrps(x)$country,
    `year/country` = x$data %>%
      distinct(country, year) %>%
      count(country) %>%
      pull(n) %>%
      median(),
    nobs_internet = standata(x)$N_internet,
    nobs_outcome = standata(x)$N_val,
    outcome_range = str_glue(
      "{number(min(x$data$val), .01)} - {number(max(x$data$val), .01)}"
    )
  )
}

fits %>%
  mutate(summary = map(fit1, msum)) %>%
  select(outcome, summary) %>%
  unnest(summary) %>%
  kbl(caption = "Model 1 data summary") %>%
  kable_custom() %>% 
  row_spec(0, bold = TRUE)
```

@@

```{r ppcheck-density}
#| column: page
#| fig-cap: "Graphical posterior predictive check by region (Model 1). Thin red lines indicate model replicates, whereas the blue lines indicate data; the model is mostly in smooth agreement with data, but has greater discrepancies in some regions (e.g. Anglo Depression)."
#| label: fig-ppcheck-1

ppcheck_1 <- fits %>%
  group_by(outcome) %>%
  mutate(
    out = map(
      fit1,
      ~ pp_check(
        .x,
        ndraws = 10,
        type = "dens_overlay_grouped",
        group = "region",
        resp = "val",
        newdata = .x$data
      )[[1]]
    )
  ) %>%
  ungroup() %>% 
  select(outcome, out) %>%
  unnest(out)

ppcheck_1 %>%
  ggplot(aes(value, col = is_y, alpha = is_y, group = rep_id)) +
  scale_color_brewer(palette = "Set1") +
  scale_alpha_manual(values = c(.2, 1)) +
  stat_density(
    geom = "line",
    size = .75,
    position = position_identity()
  ) +
  labs(x = "Value", y = "Density") +
  facet_grid2(
    outcome ~ group,
    scales = "free", independent = "x",
    labeller = as_labeller(function(x) str_replace(x, "_", "\n"))
  ) +
  theme(
    legend.position = "none",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
rm(ppcheck_1)
```

@@

```{r calculate-simple-effects}
#' Posterior draws at average and region levels
#'
#' @param x brmsfit (model 1/2)
#' @param p name of parameter
#'
#' @return a long tibble posterior draws of p for every region (including the average, so useful for e.g. forest plots)
pdraws <- function(x, p = "val_year") {
  bind_cols(
    tibble("Average" = fixef(x, summary = FALSE)[, p]),
    as_tibble(coef(x, summary = FALSE)$region[, , p])
  ) %>%
    pivot_longer(everything(), names_to = "region")
}

@@

```{r shrinkage}
#| column: screen-inset
#| fig-width: 14
#| fig-height: 9
#| fig-cap: "Scatterplots of multilevel and independent model coefficients"
#| label: fig-trend-shrinkage

@@

```{r model21-compare}
#| results: hold
#| tbl-cap: "Model 1 and 2 proportions variance explained by outcome"
#| label: tbl-m12-r2
fits %>%
  select(outcome, fit1, fit2) %>% 
  pivot_longer(c(fit1, fit2)) %>% 
  mutate(
    r2 = map(
      value,
      ~ bayes_R2(., resp = "val") %>%
        as.data.frame() %>%
        mutate(across(everything(), ~ percent(., .1))) %>%
        transmute(result = str_glue("{Estimate} [{Q2.5}, {Q97.5}]"))
    )
  ) %>%
  select(-value) %>%
  unnest(r2) %>% 
  pivot_wider(
    names_from = name,
    values_from = result,
    names_prefix = "Model "
  ) %>%
  mutate(outcome = str_replace(outcome, "_", " ")) %>%
  kbl(caption = "Model 1 and 2 R-squared values") %>%
  kable_custom()
```

@@

```{r model2-calculate-coefficients}
p_draws <- fits %>%
  mutate(pdraws = map(fit2, ~pdraws(.x, "i1_cw"))) %>%
  select(outcome, pdraws) %>%
  unnest(pdraws)

@@

```{r model2-fitted-values}
#| column: screen-inset
#| fig-width: 14
#| fig-height: 9
#| fig-cap: "Counterfactual plot of internet associations"
#| label: fig-fitted-internet

@@

```{r}
#| column: page
#| fig.width: 8
#| fig.height: 6
#| fig-cap: "Between-country associations"
#| label: fig-cb1

@@

  newy <- bind_cols(
    newx,
    as.data.frame(
      fitted(
        x, 
        newdata = newx, 
        re_formula = NA
        )
    )
  ) %>%
    select(-outcome)
  newy
}
tmp <- fits %>% 
  mutate(out = map(fit2, fitted_cb1)) %>% 
  select(outcome, out) %>%
  unnest(out)

@@

tmp %>% 
  ggplot(aes(i1_cb, val)) +
  scale_x_continuous(
    "Between-country centered internet adoption",
    labels = ~percent(.)
  ) +
  scale_y_continuous(
    "Value",
    labels = ~percent(./100)
  ) +
  geom_point(
    shape = 1, size = 1, alpha = .75
  ) +
  geom_ribbon(
    aes(ymin = Q2.5, ymax = Q97.5),
    fill = "dodgerblue2",
    alpha = .2
    ) +
  geom_line(aes(y = Estimate), col = "dodgerblue4") +
  geom_text(
    data = cb1_pars,
    aes(x = -Inf, y = Inf, label = str_glue("{mean} [{q2.5}, {q97.5}]")),
    vjust = 1.3, hjust = -0.025, family = Font
  ) +
  facet_wrap(
    "outcome",
    scales = "free_y", nrow = 2,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(
    aspect.ratio = 1,
    axis.title.y = element_blank()
  )
```

@@

```{r model1-demographics}
#| column: screen-inset
#| fig-width: 14
#| fig-height: 9
#| fig-cap: "Forest plot of linear trends in outcomes by demographic group"
#| label: fig-forest-trends-demo
h_gwp <- c(
  `15 to 19` =
    "val_year + val_year:age1 * 1 = 0",
  `20 to 24` =
    "val_year + val_year:age1 * -1 = 0",
  `Female` =
    "val_year + val_year:sex1 * 1 = 0",
  `Male` =
    "val_year + val_year:sex1 * -1 = 0"
)
h_gbd <- c(
  `10 to 14` =
    "val_year + val_year:age1 * 1 + val_year:age2 * 0 = 0",
  `15 to 19` =
    "val_year + val_year:age1 * 0 + val_year:age2 * 1 = 0",
  `20 to 24` =
    "val_year + val_year:age1 * -1 + val_year:age2 * -1 = 0",
  `Female` =
    "val_year + val_year:sex1 * 1 = 0",
  `Male` =
    "val_year + val_year:sex1 * -1 = 0"
)
#' Calculate posterior draws for simple effects (by sex and age) for regions and average
#'
#' @param x brmsfit model 2
#' @param h a brmshypothesis
#'
#' @return a long tibble of draws of posterior answers to hypothesis
simple_effects_samples <- function(x, h) {
  # Calculate hypotheses at average level
  average_h <- hypothesis(
    x, h
  )
  # Give hypotheses appropriate names
  average <- average_h$samples
  names(average) <- interaction(
    "Average", average_h$hypothesis$Hypothesis,
    sep = "+"
  )
  # Pivot to long format, note use of dtplyr
  average <- average %>%
    lazy_dt() %>%
    pivot_longer(everything()) %>%
    separate(name, into = c("region", "Hypothesis"), sep = "\\+") %>%
    as_tibble()

@@

```{r model2-demographics}
#| column: screen-inset
#| fig-width: 14
#| fig-height: 9
#| fig-cap: "Forest plot of associations by demographic group"
#| label: fig-forest-association-demo
h_gwp <- c(
  `15 to 19` =
    "i1_cw + age1:i1_cw * 1 = 0",
  `20 to 24` =
    "i1_cw + age1:i1_cw * -1 = 0",
  `Female` =
    "i1_cw + sex1:i1_cw * 1 = 0",
  `Male` =
    "i1_cw + sex1:i1_cw * -1 = 0"
)
h_gbd <- c(
  `10 to 14` =
    "i1_cw + age1:i1_cw * 1 + age2:i1_cw * 0 = 0",
  `15 to 19` =
    "i1_cw + age1:i1_cw * 0 + age2:i1_cw * 1 = 0",
  `20 to 24` =
    "i1_cw + age1:i1_cw * -1 + age2:i1_cw * -1 = 0",
  `Female` =
    "i1_cw + sex1:i1_cw * 1 = 0",
  `Male` =
    "i1_cw + sex1:i1_cw * -1 = 0"
)
